# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
unzip ("activity.zip", "activity.csv")
df <- read.csv("activity.csv")
df$date <- strptime(df$date, "%Y-%m-%d")
library(plyr)
```


## What is mean total number of steps taken per day?
```{r}
means <- ddply(df, .(date), sum=sum(steps), summarise)
barplot(means$sum)
mean_steps <- mean(means$sum[!is.na(means$sum)])
median_steps <- median(means$sum[!is.na(means$sum)])
```

Mean number of steps is `r mean_steps`, median is `r median_steps`

## What is the average daily activity pattern?
```{r}
intervals <- ddply(df, .(interval), mean=mean(steps, na.rm=T), summarise)
plot(intervals$mean, type='l', xlab="5-min interval", ylab="Avg steps")
max_idx <- which.max(intervals$mean)
```

5-minute interval with max avg steps is `r max_idx`

## Imputing missing values
```{r}
nas <- is.na (df$steps)
na_count <- sum (nas)

dfall <- data.frame(date=df$date, interval=df$interval, steps=df$steps)

for (i in which(nas)) {
  dfall$steps[i] <- intervals$mean[intervals$interval == df$interval[i]]
}

means_all <- ddply(dfall, .(date), mean=mean(steps), sum=sum(steps), summarise)
barplot(means_all$sum)
mean_all_steps <- mean(means_all$sum)
median_all_steps <- median(means_all$sum)
```

Total amount of NAs is `r na_count`

Mean number of steps for dataset without NAs is `r mean_all_steps`, median is `r median_all_steps`

## Are there differences in activity patterns between weekdays and weekends?

```{r}
wd <- weekdays(dfall$date)
we <- wd == "Saturday" | wd == "Sunday"
dfall <- data.frame(dfall, weekend=factor(we, labels=c("weekday", "weekend")))

wd_means <- ddply(dfall[dfall$weekend == "weekday",], .(interval), mean=mean(steps), summarise)
we_means <- ddply(dfall[dfall$weekend == "weekend",], .(interval), mean=mean(steps), summarise)

opar <- par(no.readonly=T)
par(mfrow=c(2,1))
plot(we_means$mean, type='l', main="weekend", xlab="5-min interval", ylab="Mean steps")
plot(wd_means$mean, type='l', main="weekday", xlab="5-min interval", ylab="Mean steps")
par(opar)
```